---
title: 'Exploratory Analysis at a Glance: Using sleep Data as an Example'
author: "Yao-Jen Kuo"
date: "Tuesday, October 06, 2015"
output:
  html_document:
    theme: united
    toc: true
    toc_depth: 2
    highlight: tango
    fig_width: 8
    fig_height: 5
    fig_caption: false
---

# Introduction

The typical procedure for a data analysis project starts with collecting data and then exploring data. In this document, we are going to introduce three common approaches to give us a basic understanding of our data: the pattern of missing values, the distribution of continuous variables, and the frequency of categorical variables.

# The Data

We are going to use the **sleep** data as our example dataset. This data originates from **Allison, T. and Chichetti, D. (1976) Sleep in mammals: ecological and constitutional correlates. Science 194 (4266), 732-734.** You can find the article from [here](http://www.faculty.biol.ttu.edu/strauss/Multivar/Lab/AllisonCicchetti1976.pdf).

The **sleep** data is a data frame with 62 observations on 10 variables. We can take a glimpse quickly:

```{r, message=FALSE, warning=FALSE}
library(VIM)
str(sleep)#The structure of sleep
summary(sleep)#The summary of sleep
```

# The Pattern of Missing Values

```{r}
aggr(sleep, combined=T, numbers=T, prop=F, main='Missing Pattern of Sleep', ylab='')#To draw the missing plot
```

It is easily observed that there are missing values in the following variables: NonD, Dream, Sleep, Span, and Gest.

# Separating Continuous Variables from Categorical Variables

```{r}
indexInt <- sapply(sleep, is.integer)#find out which variable is integer
sleepIntVar <- sleep[, indexInt]# as in categorical
sleepNumVar <- sleep[, !indexInt]#as in continuous
```

We are using histogram to explore the distribution of continuous variables and barplot to explore categorical variables.

# The Distribution of Continuous Variables

```{r, fig.height=8}
#To decide how to arrange our plots
autoMfrow <- function(x){
  if (x%%2==0){
    par(mfrow=c((x)/2,2))
  } else{
    par(mfrow=c((x+1)/2,2))
  }
}
numCnt <- ncol(sleepNumVar)
autoMfrow(numCnt)#To set the arrangement of plots
numVarName <- names(sleepNumVar)
#To plot all histograms for continuous values
for (name in numVarName){
  histNum <- hist(sleepNumVar[,name],
       freq=FALSE,
       col="skyblue",
       xlab=name,
       ylab='',
       main=paste("Histogram of: ", name))
  lines(density(sleepNumVar[,name], na.rm=TRUE), lwd=2, col="red")
}
```

# The Frequency of Categorical Variables

```{r}
intCnt <- ncol(sleepIntVar)
autoMfrow(intCnt)#To set the arrangement of plots
intVarName <- names(sleepIntVar)
#To plot all barplots for categorical values
for (name in intVarName){
  barplotInt <- barplot(table(sleepIntVar[,name]),
       xlab=name,
       col="skyblue",
       main=paste("Barplot of: ", name))
  text(x=barplotInt, y=table(sleepIntVar[,name])+2, labels=as.character(table(sleepIntVar[,name])), xpd=TRUE)
}
```